---
title: "potholes"
author: "Gabriel Florit"
output:
  html_document:
    self_contained: false
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(stringr)
library(lubridate)
#library(rgdal)
library(dplyr)
library(ggplot2)
library(scales)
library(ggmap)
#library(zoo)
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=TRUE}
rawdata <- read.csv('../potholes.csv', strip.white=TRUE, stringsAsFactors=FALSE, row.names=NULL)
```

A note on the format: each question is followed by the R code that generates the answer. This is also known as **reproducible research**, a practice that's slowly being adopted by newspapers (e.g. [538](https://github.com/fivethirtyeight/data), [The Upshot](https://github.com/theupshot)). From [wikipedia](http://en.wikipedia.org/wiki/Reproducibility#Reproducible_research): "The term reproducible research refers to the idea that the ultimate product of academic research is the paper along with the full computational environment used to produce the results in the paper such as the code, data, etc. that can be used to reproduce the results and create new work based on the research."

****
#### Before we begin: our data has `r I(nrow(rawdata))` rows and `r I(ncol(rawdata))` columns.

****
## Universal assumptions (meaning they apply to all following graphics)
- There are two "date closed" columns, **CLOSED_DT** and **Date.Closed**. Not all rows have valid entries for both columns. Since we need one master "date closed" column, we'll have to use **CLOSED_DT** unless blank, in which case we'll use **Date.Closed**.
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE, cache=TRUE}
# Get a list of rows with valid CLOSED_DT data,
# and use it to create a new column, DATE.CLOSED.R.
closed_dt_rows <- rawdata %>%
  filter(nchar(CLOSED_DT) > 0) %>%
  mutate(DATE.CLOSED.R = mdy(CLOSED_DT))

# Get a list of rows with invalid CLOSED_DT data,
# and use Date.Closed to create a new column, DATE.CLOSED.R.
date.closed_rows <- rawdata %>%
  filter(nchar(CLOSED_DT) == 0) %>%
  mutate(DATE.CLOSED.R = mdy(Date.Closed))

# Merge all rows together.
potholes <- rbind(
  closed_dt_rows,
  date.closed_rows
)
```

****
#### How many potholes were closed per year?
```{r, results='asis', fig.width=10, fig.height=6, warning=FALSE, message=FALSE}
potholes %>%
  filter(!is.na(DATE.CLOSED.R)) %>%
  mutate(YEAR = year(DATE.CLOSED.R)) %>%
  group_by(YEAR) %>%
  summarise(POTHOLES = n()) %>%
  knitr::kable()
```

****
#### What does this look like over time?
```{r, results='asis', out.width='910px', fig.width=10, fig.height=2, warning=FALSE, message=FALSE, dpi=50}
data <- potholes %>%
  filter(!is.na(DATE.CLOSED.R)) %>%
  arrange(DATE.CLOSED.R) %>%
  group_by(DATE.CLOSED.R) %>%
  summarise(closures = n())

write.csv(data, file='output/potholeClosuresPerDay.csv', row.names=FALSE)

ggplot(data, aes(DATE.CLOSED.R, closures)) +
  geom_bar(stat='identity') +
  ggtitle('Daily pothole tally') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

****
#### Which districts are contributing to the increase in pothole closures?
```{r, results='asis', out.width='910px', fig.width=10, fig.height=2, warning=FALSE, message=FALSE, dpi=50}
data <- rbind(
  (potholes %>%
    filter(pwd_district == '2') %>%
    mutate(district = 'District 2')),
  (potholes %>%
    filter(pwd_district != '2') %>%
    mutate(district = 'Rest'))) %>%
  mutate(YEAR = year(DATE.CLOSED.R)) %>%
  filter(!is.na(YEAR)) %>%
  group_by(district, YEAR) %>%
  tally() %>%
  summarise(increase = diff(n))

write.csv(data, 'output/yearlyIncreaseByDistrict.csv', row.names = FALSE)

ggplot(data, aes(x=district, y=increase, fill=district)) +
  geom_bar(stat='identity') +
  geom_text(aes(label=increase), hjust=1) +
  ylab('potholes fixed') +
  xlab(NULL) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  coord_flip() +
  ggtitle('2014 increase over 2013, by district')
```

****
#### Let's look at district 2 potholes filled per week.
```{r, results='asis', out.width='910px', fig.width=10, fig.height=2, warning=FALSE, message=FALSE, dpi=50}
data <- potholes %>%
  filter(
    !is.na(DATE.CLOSED.R),
    pwd_district == '2'
  ) %>%
  transmute(WEEK = floor_date(DATE.CLOSED.R, 'week')) %>%
  arrange(WEEK) %>%
  group_by(WEEK) %>%
  tally()

write.csv(data, 'output/weeklyClosuresForDistrict2.csv', row.names = FALSE)

ggplot(data, aes(WEEK, n)) +
  geom_bar(stat='identity') +
  ggtitle('Weekly pothole closures for district 2') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

****
#### Let's look at District 2's top day of filling potholes.
## Graphic-specific assumptions
- We will ignore potholes located at (-71.0587,42.3594), which is City Hall. 
```{r, results='asis', out.width='910px', fig.width=10, fig.height=7, warning=FALSE, message=FALSE, dpi=50, dev='jpeg'}
district <- 2

data <- potholes %>%
  filter(
    !is.na(DATE.CLOSED.R),
    pwd_district == district,
    LONGITUDE != -71.0587,
    LATITUDE != 42.3594
  ) %>%
  arrange(DATE.CLOSED.R) %>%
  group_by(DATE.CLOSED.R, pwd_district) %>%
  tally() %>%
  group_by(pwd_district) %>%
  slice(which.max(n)) %>%
  inner_join(potholes, by=c('DATE.CLOSED.R', 'pwd_district')) %>%
  filter(
    LONGITUDE != -71.0587,
    LATITUDE != 42.3594
  ) %>%
  select(
    LATITUDE,
    LONGITUDE,
    DATE.CLOSED.R
  ) %>%
  group_by(LATITUDE, LONGITUDE, DATE.CLOSED.R) %>%
  tally() %>%
  ungroup() %>%
  arrange(desc(n))

csv <- data %>%
  select(LATITUDE,LONGITUDE,n)

csv <- cbind(row = rownames(csv), csv)

write.csv(csv, str_c('output/bestDayForDistrict', district, '_', unique(data$DATE.CLOSED.R), '.csv'), row.names = FALSE)

map <- get_map(location=c(min(data$LONGITUDE), min(data$LATITUDE), max(data$LONGITUDE), max(data$LATITUDE)), zoom=13)
ggmap(map) +
  geom_point(aes(x=LONGITUDE, y=LATITUDE, size=n, alpha=0.05), data=data) +
  geom_point(shape=1, aes(x=LONGITUDE, y=LATITUDE, size=n), data=data) +
  scale_size_area(max_size=20) +
  ggtitle(str_c('District ', district, ' fixed ', sum(data$n), ' potholes in ', nrow(data) , ' locations on ', unique(data$DATE.CLOSED.R)))
```

****
#### Show me clusters of 15 or more potholes fixed on the same day for all districts for 2014.
## Graphic-specific assumptions
- We will ignore potholes located at (-71.0587,42.3594), which is City Hall. 
```{r, results='asis', out.width='910px', fig.width=10, fig.height=7, warning=FALSE, message=FALSE, dpi=50, dev='jpeg'}
data <- rbind(
  (potholes %>%
    filter(pwd_district %in% c('2', '3')) %>%
    mutate(district = pwd_district)),
  (potholes %>%
    filter(!(pwd_district %in% c('2', '3'))) %>%
    mutate(district = 'Rest'))) %>%
  select(-pwd_district) %>%
  filter(
    LATITUDE!=42.3594, LONGITUDE!=-71.0587,
    year(DATE.CLOSED.R) == 2014
  ) %>%
  group_by(DATE.CLOSED.R, LONGITUDE, LATITUDE, district) %>%
  tally() %>%
  ungroup() %>%
  filter(n >= 15) %>%
  select(-DATE.CLOSED.R) %>%
  arrange(desc(n))

csv <- cbind(row = rownames(data), data)

write.csv(csv, 'output/clustersIn2014.csv', row.names = FALSE)

map <- get_map(location=c(min(data$LONGITUDE), min(data$LATITUDE), max(data$LONGITUDE), max(data$LATITUDE)), zoom=12)
ggmap(map) +
  geom_point(aes(x=LONGITUDE, y=LATITUDE, size=n, alpha=0.05,color=district), data=data) +
  geom_point(shape=1, aes(x=LONGITUDE, y=LATITUDE, size=n), data=data) +
  ggtitle(str_c(nrow(data), ' single-day clusters')) +
  scale_size_area(max_size=15)
```

```{r, results='asis', out.width='910px', fig.width=10, fig.height=2, warning=FALSE, message=FALSE, dpi=50}
clusters <- data %>%
  group_by(district) %>%
  summarise(count = n())

ggplot(clusters, aes(x=district, y=count, fill=district)) +
  geom_bar(stat='identity') +
  geom_text(aes(label=count), hjust=1) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  coord_flip() +
  ggtitle(str_c(nrow(data), ' single-day clusters'))
```

****
#### In the above map, what district fixed the cluster near Chinatown, and did they took care of the same location previously?
```{r, results='asis', out.width='910px', fig.width=10, fig.height=2, warning=FALSE, message=FALSE, dpi=50}
potholes %>%
  filter(LATITUDE == 42.3520, LONGITUDE == -71.0615) %>%
  group_by(LATITUDE, LONGITUDE, pwd_district, DATE.CLOSED.R) %>%
  tally() %>%
  knitr::kable()
```
